Structured Versus Decorated Cospans

### 2022-02-18 ResponseToReviews.tex

pg. 3: “Many of the flawed applications of decorated cospans have been fixed using structured cospans.” This dismisses the work of other scholars 
using the decorated cospan approach as flawed. The alleged flaw should be made more precise. What about the decorated cospan approach is inferior to 
the structured cospan approach? The goal of this paper should be to illustrate to an applied category theory audience the differences between 
decorated and structured cospans. Referring back to previous work of the authors to make that point does not provide value.

pg. 27: To show that U does not have a left adjoint, we use the following well-known result. For readers of the TAC paper previously cited, that might 
be a well known result that does not require a citation, but my understanding of the Compositionality audience assumed that the ACT work was targeted 
readers that are more familiar with the applications and thus cannot be assumed to know such results.}

\begin{verbatim}
[1] https://arxiv.org/abs/2106.04703
[2] https://github.com/AlgebraicJulia/Catlab.jl/blob/f06417de29421441c4133c1cfb75b39ae8aa7142/src/categorical_algebra/StructuredCospans.jl
[3] https://github.com/AlgebraicJulia/AlgebraicPetri.jl
\end{verbatim}

\subsection{Third review}

{\footnotesize
Decorated and structured cospans have emerged as a mainstay of the current wave of ACT, the latter correcting certain technical problems with the 
former. This paper poses two questions:

    How can the technical problems with decorated cospans be fixed?

    How does the corrected version of decorated cospans compare with structured cospans?

The paper offers the following answers:

The symmetric lax monoidal functor (A,+) -> (Set, $\times$) defining a category of isomorphism classes of decorated cospans is replaced by a 
symmetric lax monoidal pseudofunctor (A,+) -> (Cat, $\times$) defining a symmetric monoidal double category of decorated cospans.

A sufficient condition, the main Theorem 4.1, is provided for a double category of decorated cospans to be equivalent, indeed isomorphic, to a double 
category of structured cospans. Several examples, as well as a nonexample, of applying the theorem are given.

The paper is a technical one, giving slick but abstract constructions and proofs using machinery developed by Shulman for turning monoidal 
bifibrations into double categories, and it serves more to clarify and compare existing ACT methods than to introduce new ones. The mathematical 
exposition is, however, very strong, and a fairly wide range of readers should be able to take something away from the paper. Given the central status 
of decorated/structured cospans within ACT, it was important that the above questions be addressed, and the authors have done so admirably. Also, the 
constructions and results are “black-boxed” in such a manner that practioners will be able to use them without understanding in detail the proofs that 
support them. I therefore consider this paper to be a strong contribution to the scholarship of ACT and recommend that it be published.

Following are some questions and minor suggested changes.

p. 2. “Unfortunately, many applications of decorated cospans were flawed… for reasons painstaking explained in [1]”

Since fixing these problems is a major motivation for the whole paper, it would be good to explain what they are in slightly more detail, rather than 
just giving a reference. This could be done in a one paragraph or less. The gist of it, as I recall, is that in the original approach to decorated 
cospans, the isomorphism classes of open graphs depend on the identities of the edge elements, which is clearly wrong.

p. 4, following statement of Thm 2.1.

The diagram depicting the decoration morphism $\tau$ as a natural transformation, first shown explictly on p. 10, should be moved here since it is 
implicitly used at the beginning of p. 5 to define the decoration of horizontal composites.

p. 9, ingredients for constructing a double functor of decorated cospan double categories.

Does the natural transformation $\theta$: EF => F’ H have to be a natural isomorphism? Quite possibly I missed something, but I don’t see where in 
the following proofs that assumption is used.

p. 21, definition of F: Set -> Cat

The use f, g for vertex/edge maps is opposite of the convention used on the previous page. I suggest that the same convention be used for both the 
decorated and structured approach to constructing open graphs.

p. 27, end of Section 6.

It seems that the left adjoint needed for structured cospans fails to exist because the chosen category of dynamical systems does not have enough 
morphisms. Relatedly, this is the only example where the pseudofunctor D: (FinSet,+) -> (Cat, $\times$) comes from a simple functor D: (FinSet,+) -> 
(Set, $\times$) by viewing sets as discrete categories. In the other examples, this would count as using decorated cospans in the old, flawed way. 
All of this suggests the question: is there a way of constructing a category of dynamical systems that has more morphisms, so that a connection with 
structured cospans can be established?

p. 30, following Thm A.2

If I am not mistaken, “fibration”, “opfibration,” “monoidal (op/bi)fibration” are all defined, but the definition of “bifibration” is omitted. For 
completeness, it should be included.

A last general comment is that the paper unfortunately adopts the strategy, all too common in math papers, of putting all the examples in a section at 
the end of the paper. I think the readability of the paper would be improved, without much change to the organization, if the motivating example of 
open graphs were added to the relevant places in Sections 2 and 3. The remaining examples could be left in Section 6.
}

\subsection{Fourth review}

{\footnotesize
What are the main questions posed by the manuscript, and how does it answer them? Why or why not is this work a significant technical or conceptual 
contribution to scholarship?

Decorated cospans and structured cospans are two different tools for building categories of open systems, both of which have been applied in several 
existing papers (and this reviewer believes they will continue to grow in importance over time). This paper explains the relationship between them, in 
a very elegant way using the equivalence between fibered and indexed categories, to the extent that the two constructions now appear as different 
perspectives on the same fundamental idea. To me this relationship appears obvious in retrospect, and I mean this as a compliment to the paper. I 
think this paper is an important contribution to categorical systems theory, and I recommend acceptance.

Composition of morphisms f, g in this paper is usually denoted gf, and occasionally gof. I’d suggest mentioning this in the “conventions” section.

p3: “Using the Grothendieck construction” - maybe worth mentioning this is the “covariant” version which is different to the most common convention 
(at least I think so, but the literature seems to be very split over this)

Theorem 2.1: Maybe worth saying explicitly that s is an object of the category F(m) rather than using $\in$, since this is exactly the distinction to 
Fong’s original definition

Proof of theorem 2.1: You should probably at least handwave that the cartesian liftings really are cartesian liftings. (No doubt it’s easy, but this 
is a classic sort of spot where errors can hide)

On page 9 during the construction of H, I would appreciate a sentence giving some intuition for the functor E, at this point I’m confused about its 
role.

Theorem 2.5: F should be defined on A, not A’

I think it would be worth saying in words a bit more prominently what the hypothesis of theorem 4.1 roughly means, namely that the fibers of the total 
category have finite colimits, reindexing preserves them, and the monoidal product on the fibers is cocartesian.

I wonder if the “problem” where DCsp is not a structured cospan category could be “fixed” by upgrading the functor from Set to Cat in a non-discrete 
way, for example with an ordering with the zero vector field as the bottom element. I suppose this would give you different 2-cells in the resulting 
double category or bicategory, but perhaps it would yield the same 1-category}

@@

\item pg. 27: To show that U does not have a left adjoint, we use the following well-known result. For readers of the TAC paper previously cited, 
that might 
be a well known result that does not require a citation, but my understanding of the Compositionality audience assumed that the ACT work was targeted 
readers that are more familiar with the applications and thus cannot be assumed to know such results.

{\john We can just leave out "well-known".}
{\kenny Emily Riehl's Category Theory In Context (Lemma 4.6.1 on pdf page 162):

was where I found that theorem when I was working on how to show that counterexample. I doubt she was the first, but she states it exactly as we have 
it and she gives a proof.}

{\chris Third starts here}
\item p. 2. “Unfortunately, many applications of decorated cospans were flawed… for reasons painstaking explained in [1]”

Since fixing these problems is a major motivation for the whole paper, it would be good to explain what they are in slightly more detail, rather than 
just giving a reference. This could be done in a one paragraph or less. The gist of it, as I recall, is that in the original approach to decorated 
cospans, the isomorphism classes of open graphs depend on the identities of the edge elements, which is clearly wrong.

\item p. 4, following statement of Thm 2.1.

The diagram depicting the decoration morphism $\tau$ as a natural transformation, first shown explictly on p. 10, should be moved here since it is 
implicitly used at the beginning of p. 5 to define the decoration of horizontal composites.

\item p. 9, ingredients for constructing a double functor of decorated cospan double categories.

Does the natural transformation $\theta$: EF => F’ H have to be a natural isomorphism? Quite possibly I missed something, but I don’t see where in 
the following proofs that assumption is used.

\item p. 21, definition of F: Set -> Cat

The use f, g for vertex/edge maps is opposite of the convention used on the previous page. I suggest that the same convention be used for both the 
decorated and structured approach to constructing open graphs.

\item p. 27, end of Section 6.

It seems that the left adjoint needed for structured cospans fails to exist because the chosen category of dynamical systems does not have enough 
morphisms. Relatedly, this is the only example where the pseudofunctor D: (FinSet,+) -> (Cat, $\times$) comes from a simple functor D: (FinSet,+) -> 
(Set, $\times$) by viewing sets as discrete categories. In the other examples, this would count as using decorated cospans in the old, flawed way. 
All of this suggests the question: is there a way of constructing a category of dynamical systems that has more morphisms, so that a connection with 
structured cospans can be established?

\item p. 30, following Thm A.2

If I am not mistaken, “fibration”, “opfibration,” “monoidal (op/bi)fibration” are all defined, but the definition of “bifibration” is omitted. For 
completeness, it should be included.

{\john If we don't define bifibration, we definitely should!}

\item A last general comment is that the paper unfortunately adopts the strategy, all too common in math papers, of putting all the examples in a 
section at 
the end of the paper. I think the readability of the paper would be improved, without much change to the organization, if the motivating example of 
open graphs were added to the relevant places in Sections 2 and 3. The remaining examples could be left in Section 6.

{\john
I don't want to bust up the examples and put some in sections 2 and 3. Anyone who wants examples can easily find them all in Section 6, we mention 
this section in the introduction (and perhaps should advertise it more there), and it's hard to analyze the examples without having all the theory in 
hand.}

{\chris Fourth starts here}
\item Composition of morphisms f, g in this paper is usually denoted gf, and occasionally gof. I’d suggest mentioning this in the “conventions” 
section.

\item p3: “Using the Grothendieck construction” - maybe worth mentioning this is the “covariant” version which is different to the most common 
convention 
(at least I think so, but the literature seems to be very split over this)

\item Theorem 2.1: Maybe worth saying explicitly that s is an object of the category F(m) rather than using $\in$, since this is exactly the 
distinction to 
Fong’s original definition

\item Proof of theorem 2.1: You should probably at least handwave that the cartesian liftings really are cartesian liftings. (No doubt it’s easy, but 
this 
is a classic sort of spot where errors can hide)

\item On page 9 during the construction of H, I would appreciate a sentence giving some intuition for the functor E, at this point I’m confused about 
its 
role.

\item Theorem 2.5: F should be defined on A, not A’

\item I think it would be worth saying in words a bit more prominently what the hypothesis of theorem 4.1 roughly means, namely that the fibers of 
the total 
category have finite colimits, reindexing preserves them, and the monoidal product on the fibers is cocartesian.

\item I wonder if the “problem” where DCsp is not a structured cospan category could be “fixed” by upgrading the functor from Set to Cat in a 
non-discrete 
way, for example with an ordering with the zero vector field as the bottom element. I suppose this would give you different 2-cells in the resulting 
double category or bicategory, but perhaps it would yield the same 1-category
\end{enumerate}

### 2022-03-04 cospans_new.tex
first round of easy edits

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $(F,\phi,\phi_0) \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $s \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a \define{decoration morphism} $\tau \maps F(h)(s) \to s'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

\begin{lem}
\label{lem:MonGroth}
For any monoidal category $(\A,\otimes)$, there is a 2-equivalence between the 2-categories of monoidal opfibrations $U\maps (\X, \otimes) \to (\A, \otimes)$ and monoidal pseudofunctors $F \maps (\A,\otimes) \to (\mathsf{Cat}, \times)$, and if $\A$ cocartesian monoidal, there is a 2-equivalence between these and pseudofunctors from $\A$ into $\MonCat$.  If $\A$ is symmetric monoidal, there is also a 2-equivalence between the 2-categories of symmetric monoidal opfibrations $U \maps (\X, \otimes) \to (\A, \otimes)$ and symmetric monoidal pseudofunctors $F \maps  (\A,\otimes) \to (\mathsf{Cat}, \times)$ , and if $\A$ cocartesian monoidal, there is a 2-equivalence between these and pseudofunctors from $\A$ into $\SMC$.
\end{lem}

@@

where the decoration morphism $\lH(\tau)$ is defined as follows.  We can think of $\tau$ as a natural transformation
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(m)$};
\node (I) at (-2,-1) {$F(n)$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
(G) edge node [above] {$s$} (H)
(G) edge node [below] {$s'$} (I)
(H) edge node [right] {$F(h)$} (I);
and in the same style of notation $\lH(\tau)$ is given by

@@

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a monoidal natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\scriptstyle\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\lH \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}

@@

\begin{thm} \label{thm:equiv}
Suppose $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor. If the corresponding pseudofunctor $F \maps \A \to \SMC$ from \cref{lem:MonGroth}
 factors through $\Rex$, then the symmetric monoidal double categories $F\lCsp$ of decorated cospans and ${}_L\lCsp(\inta F)$ of structured cospans are isomorphic, where $L \maps \A \to \inta F$ is a left adjoint of the induced Grothendieck opfibration $U \maps \inta F \to \A$.
\end{thm}

@@

@@

\begin{cor} \label{cor:fcocMonGroth}
 Suppose $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor for which the corresponding pseudofunctor $F  \maps \A \to \SMC$ from \cref{lem:MonGroth} factors through $\Rex$.  Then $\inta F$ has all finite colimits and  the induced opfibration $U\maps \inta F \to \A$ preserves them.  Moreover we can choose finite colimits in $\inta F$ so that $U$ preserves them strictly.
\end{cor}

Regarding the assumptions of the above corollary, notice that since $\A$ is cocartesian monoidal, the lax monoidal pseudofunc\-tor structure $(F,\phi,\phi_0)$ gives rise to a specific symmetric monoidal structure on the fibers $F(a)$ in terms of $\phi,\phi_0$, given explicitly by \cref{eq:explicitstructure1}. Since we now ask that the pseudofunctor $F\maps \A\to\SMC$ factors through $\Rex$, this fiberwise monoidal structure is required to be cocartesian, namely that \cref{eq:explicitstructure1} gives coproducts and an initial object in each $F(a)$.

@@

\begin{proof}
 This is \cite[Proposition~4.4]{Gray}.    Suppose each fiber $\X_a$ of the opfibration $U$ has an initial object $\bot_a$ and these objects are preserved (up to isomorphism) by the reindexing functors.   Define $L \maps \A \to \X$ on objects $a \in \A$ by $L(a) = \bot_a$.   Given a morphism $f\maps a\to a'$ in $\A$, define $L(f)$ to be the composite
\[
  \bot_a\xrightarrow{\mathrm{Cocart}(f,\bot_a)}f_!(\bot_a)\xrightarrow{\;\chi_a\;}\bot_{a'}
\]
where $\mathrm{Cocart}(f,\bot_a)$ is the cocartesian lifting of $f$ to $\bot_a$ and $\chi_a$ is the unique isomorphism between two initial objects in the fiber above $a'$.    The functor $L$ then becomes left adjoint to $R$ with unit $\iota_a \maps a \to U(L(a))$ being the identity, using the fact that $U(L(a)) = U(\bot_a) = a$.
\end{proof}

@@

All the above concepts and results have analogues for fibrations.   A functor $\Phi \maps \X \to \A$ a \define{fibration} if and only if $\Phi\op \maps \X\op \to \A\op$ is an opfibration.  Equivalently, $\Phi$ is a fibration if and only if for every $y \in \X$ with $U(y) = b$ and $f \maps a \to b$ in $\A$ there exists a cartesian lifting of $f$ to $y$, where this concept is defined dually to cocartesian lifting.     Furthermore, if $\Phi$ is a fibration there is a contravariant reindexing functor
\[    f^\ast \maps \X_b \to \X_a \]
for each morphism $f \maps a \to b$.

@@

In a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f_!$ is left adjoint to $f^*$ \cite[Proposition 3.9]{Shulman2008}.   Using this, one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a specified natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\C_a$};
\node (B) at (1.5,0) {$\C_b$};
\node (C) at (0,-1) {$\C_c$};
\node (D) at (1.5,-1) {$\C_d$};
\node (E) at (0.75,-0.5) {$\scriptstyle\Downarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
where $\theta$ is built as a composite involving the unit of the adjunction between
$g_!$ and $g^\ast$ and the counit of the adjunction between $k_!$ and $k^\ast$:
\begin{equation}
\label{theta}
   k_! h^\ast \To k_! h^\ast g^\ast g_! \cong k_! k^\ast f^\ast g_! \To f^\ast g_! . 
\end{equation}
If $\theta$ is a natural isomorphism whenever the original square in $\A$ is
a pushout, we say that the bifibration $\Phi$ is \define{Beck--Chevalley}.  (Shulman  uses the term `strongly co-BC' \cite[Definition 13.21] {Shulman2008}.)

>>>>

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $(F,\phi,\phi_0) \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $s \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a \define{decoration morphism} $\tau \maps F(h)(s) \to s'$ in $F(m')$, which can be thought of as a natural transformation
\begin{equation}\label{eq:taunattransf}
\begin{tikzpicture}[baseline=(current  bounding  box.center),scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(m)$};
\node (I) at (-2,-1) {$F(n)$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node [above] {$s$} (H)
(G) edge node [below] {$s'$} (I)
(H) edge node [right] {$F(h)$} (I);
\end{tikzpicture}
\end{equation}
\end{itemize}
\end{thm}

@@

\begin{lem}
\label{lem:MonGroth}
For any monoidal category $(\A,\otimes)$, there is a 2-equivalence between the 2-categories of monoidal opfibrations $U\maps (\X, \otimes) \to (\A, 
\otimes)$ and lax monoidal pseudofunctors $F \maps (\A,\otimes) \to (\mathsf{Cat}, \times)$, and if $\A$ cocartesian monoidal, there is a 
2-equivalence between these and pseudofunctors from $\A$ into $\MonCat$.  If $\A$ is symmetric monoidal, there is also a 2-equivalence between the 
2-categories of symmetric monoidal opfibrations $U \maps (\X, \otimes) \to (\A, \otimes)$ and symmetric lax monoidal pseudofunctors $F \maps  
(\A,\otimes) \to (\mathsf{Cat}, \times)$ , and if $\A$ cocartesian monoidal, there is a 2-equivalence between these and pseudofunctors from $\A$ into 
$\SMC$.
\end{lem}

@@

where the decoration morphism $\lH(\tau)$ is defined as follows.  Thinking of $\tau$ as a natural transformation \cref{eq:taunattransf},
in the same style of notation $\lH(\tau)$ is given by

@@

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, \maps (\A,+) \to (\Cat,\times)$ and $F' \maps (\A',+) \to 
(\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and 
a monoidal natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\scriptstyle\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\lH \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}

@@

\begin{thm} \label{thm:equiv}
Suppose $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor. If the corresponding pseudofunctor $F \maps \A \to \SMC$ from \cref{lem:MonGroth}
 factors through $\Rex\to\SMC$, then the symmetric monoidal double categories $F\lCsp$ of decorated cospans and ${}_L\lCsp(\inta F)$ of structured 
cospans are isomorphic, where $L \maps \A \to \inta F$ is a left adjoint of the induced Grothendieck opfibration $U \maps \inta F \to \A$.
\end{thm}

@@

Essentially, the above assumptions state that the fibers of the induced opfibration $U$ have finite colimits which are preserved by the reindexing 
functors, and the induced fiberwise monoidal structure is cocartesian. Explicitly, regarding the last point, the lax monoidal 
pseudofunctor structure $(F,\phi,\phi_0)$ gives rise to a specific symmetric monoidal structure on the fibers $F(a)$ in terms of $\phi,\phi_0$ given 
by \cref{eq:explicitstructure1}, since $\A$ is cocartesian monoidal. We here ask that the induced pseudofunctor $F\maps \A\to\SMC$ factors through 
$\Rex$, namely that \cref{eq:explicitstructure1} gives coproducts and an initial object in each $F(a)$.

@@

\begin{cor} \label{cor:fcocMonGroth}
 Suppose $\A$ has finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ is a symmetric lax monoidal pseudofunctor for which the corresponding 
pseudofunctor $F  \maps \A \to \SMC$ from \cref{lem:MonGroth} factors through $\Rex\to\SMC$.  Then $\inta F$ has all finite colimits and  the induced 
opfibration $U\maps \inta F \to \A$ preserves them.  Moreover we can choose finite colimits in $\inta F$ so that $U$ preserves them strictly.
\end{cor}

@@

\begin{proof}
 This is \cite[Proposition~4.4]{Gray}.    Suppose each fiber $\X_a$ of the opfibration $U$ has an initial object $\bot_a$ and these objects are preserved (up to isomorphism) by the reindexing functors.   Define $L \maps \A \to \X$ on objects $a \in \A$ by $L(a) = \bot_a$.   Given a morphism $f\maps a\to a'$ in $\A$, define $L(f)$ to be the composite
\[
  \bot_a\xrightarrow{\mathrm{Cocart}(f,\bot_a)}f_!(\bot_a)\xrightarrow{\;\chi_a\;}\bot_{a'}
\]
where $\mathrm{Cocart}(f,\bot_a)$ is the cocartesian lifting of $f$ to $\bot_a$ and $\chi_a$ is the unique isomorphism between two initial objects in 
the fiber above $a'$.    The functor $L$ then becomes left adjoint to $L$ with unit $\iota_a \maps a \to U(L(a))$ being the identity, using the fact 
that $U(L(a)) = U(\bot_a) = a$.
\end{proof}

@@

All the above concepts and results have analogues for fibrations.   A functor $\Phi \maps \X \to \A$ a \define{fibration} if and only if $\Phi\op \maps \X\op \to \A\op$ is an opfibration.  Equivalently, $\Phi$ is a fibration if and only if for every $y \in \X$ with $U(y) = b$ and $f \maps a \to b$ in $\A$ there exists a cartesian lifting of $f$ to $y$, where this concept is defined dually to cocartesian lifting.     Furthermore, if $\Phi$ is a fibration there is a contravariant reindexing functor
\[    f^\ast \maps \X_b \to \X_a \]
for each morphism $f \maps a \to b$.  Moreover, $\Phi\colon\X\to\A$ is a \define{bifibration} if it is both a fibration and opfibration.

@@

In a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f_!$ is left adjoint to $f^*$ \cite[Proposition 3.9]{Shulman2008}.   Using this, one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a specified natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\X_a$};
\node (B) at (1.5,0) {$\X_b$};
\node (C) at (0,-1) {$\X_c$};
\node (D) at (1.5,-1) {$\X_d$};
\node (E) at (0.75,-0.5) {$\scriptstyle\Downarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
where $\theta$ is built as a composite involving the unit of the adjunction between
$g_!$ and $g^\ast$ and the counit of the adjunction between $k_!$ and $k^\ast$:
\begin{equation}
\label{theta}
   k_! h^\ast \To k_! h^\ast g^\ast g_! \cong k_! k^\ast f^\ast g_! \To f^\ast g_! . 
\end{equation}
If $\theta$ is a natural isomorphism whenever the original square in $\A$ is
a pushout, we say that the bifibration $\Phi$ is \define{Beck--Chevalley}.  (Shulman  uses the term `strongly co-BC' \cite[Definition 13.21] {Shulman2008}.)
